\documentclass[10pt]{memoir}

% Page geometry: 4.25in x 11in (8.5x11 folded along long axis)
\usepackage[
  paperwidth=4.25in,
  paperheight=11in,
  top=0.5in,
  bottom=0.5in,
  inner=0.5in,
  outer=0.4in
]{geometry}

\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage{verse}        % better verse environment
\usepackage{setspace}
\usepackage{parskip}

% Optional: a nicer serif font (comment out if not installed)
% \usepackage{ebgaramond}

% Suppress chapter/section numbers
\setcounter{secnumdepth}{0}

% Memoir tweaks for a small poetry booklet
\pagestyle{plain}
\setlength{\parindent}{0pt}

% Verse environment settings
\setlength{\vleftmargin}{0em}   % no extra indent on verse blocks

\newcommand{\poemauthor}[1]{{\small\itshape #1}\par\medskip}

\title{Poems}

\begin{document}

\frontmatter
\maketitle
\clearpage

\mainmatter

% ── Poem 1 ──────────────────────────────────────────────────
\poemtitle{Ulysses}
\poemauthor{Alfred Lord Tennyson}

\begin{verse}
It little profits that an idle king,\\
By this still hearth, among these barren crags,\\
Match'd with an aged wife, I mete and dole\\
Unequal laws unto a savage race,\\
That hoard, and sleep, and feed, and know not me.\\
I cannot rest from travel: I will drink\\
Life to the lees: All times I have enjoy'd\\
Greatly, have suffer'd greatly, both with those\\
That loved me, and alone, on shore, and when\\
Thro' scudding drifts the rainy Hyades\\
Vext the dim sea: I am become a name;\\
For always roaming with a hungry heart\\
Much have I seen and known; cities of men\\
And manners, climates, councils, governments,\\
Myself not least, but honour'd of them all;\\
And drunk delight of battle with my peers,\\
Far on the ringing plains of windy Troy.\\
I am a part of all that I have met;\\
Yet all experience is an arch wherethro'\\
Gleams that untravell'd world whose margin fades\\
For ever and forever when I move.\\
How dull it is to pause, to make an end,\\
To rust unburnish'd, not to shine in use!\\
As tho' to breathe were life! Life piled on life\\
Were all too little, and of one to me\\
Little remains: but every hour is saved\\
From that eternal silence, something more,\\
A bringer of new things; and vile it were\\
For some three suns to store and hoard myself,\\
And this gray spirit yearning in desire\\
To follow knowledge like a sinking star,\\
Beyond the utmost bound of human thought.\\!

\vin         This is my son, mine own Telemachus,\\
To whom I leave the sceptre and the isle,—\\
Well-loved of me, discerning to fulfil\\
This labour, by slow prudence to make mild\\
A rugged people, and thro' soft degrees\\
Subdue them to the useful and the good.\\
Most blameless is he, centred in the sphere\\
Of common duties, decent not to fail\\
In offices of tenderness, and pay\\
Meet adoration to my household gods,\\
When I am gone. He works his work, I mine.\\!

\vin         There lies the port; the vessel puffs her sail:\\
There gloom the dark, broad seas. My mariners,\\
Souls that have toil'd, and wrought, and thought with me—\\
That ever with a frolic welcome took\\
The thunder and the sunshine, and opposed\\
Free hearts, free foreheads—you and I are old;\\
Old age hath yet his honour and his toil;\\
Death closes all: but something ere the end,\\
Some work of noble note, may yet be done,\\
Not unbecoming men that strove with Gods.\\
The lights begin to twinkle from the rocks:\\
The long day wanes: the slow moon climbs: the deep\\
Moans round with many voices. Come, my friends,\\
'T is not too late to seek a newer world.\\
Push off, and sitting well in order smite\\
The sounding furrows; for my purpose holds\\
To sail beyond the sunset, and the baths\\
Of all the western stars, until I die.\\
It may be that the gulfs will wash us down:\\
It may be we shall touch the Happy Isles,\\
And see the great Achilles, whom we knew.\\
Tho' much is taken, much abides; and tho'\\
We are not now that strength which in old days\\
Moved earth and heaven, that which we are, we are;\\
One equal temper of heroic hearts,\\
Made weak by time and fate, but strong in will\\
To strive, to seek, to find, and not to yield.\\!

\end{verse}

\clearpage

% ── Poem 2 ──────────────────────────────────────────────────
\poemtitle{Childe Roland to the Dark Tower Came}
\poemauthor{Robert Browning}
{\small\itshape (See Edgar's song in ``Lear.'')\par}\medskip

\begin{verse}
1.\\
My first thought was, he lied in every word,\\
\vin That hoary cripple, with malicious eye\\
\vin Askance to watch the working of his lie\\
On mine, and mouth scarce able to afford\\
Suppression of the glee that pursed and scored\\
\vin Its edge at one more victim gained thereby.\\!

2.\\
What else should he be set for, with his staff?\\
\vin What, save to waylay with his lies, ensnare\\
\vin All travellers who might find him posted there,\\
And ask the road? I guessed what skull-like laugh\\
Would break, what crutch 'gin write my epitaph\\
\vin For pastime in the dusty thoroughfare,\\!

3.\\
If at his counsel I should turn aside\\
\vin Into that ominous tract which, all agree,\\
\vin Hides the Dark Tower. Yet acquiescingly\\
I did turn as he pointed; neither pride\\
Nor hope rekindling at the end descried,\\
\vin So much as gladness that some end might be.\\!

4.\\
For, what with my whole world-wide wandering,\\
\vin What with my search drawn out thro' years, my hope\\
\vin Dwindled into a ghost not fit to cope\\
With that obstreperous joy success would bring,---\\
I hardly tried now to rebuke the spring\\
\vin My heart made, finding failure in its scope.\\!

5.\\
As when a sick man very near to death\\
\vin Seems dead indeed, and feels begin and end\\
\vin The tears and takes the farewell of each friend,\\
And hears one bid the other go, draw breath\\
Freelier outside, (``since all is o'er'' he saith,\\
\vin ``And the blow fall'n no grieving can amend'')\\!

6.\\
While some discuss if near the other graves\\
\vin Be room enough for this, and when a day\\
\vin Suits best for carrying the corpse away,\\
With care about the banners, scarves and staves,---\\
And still the man hears all, and only craves\\
\vin He may not shame such tender love and stay.\\!

7.\\
Thus, I had so long suffered in this quest,\\
\vin Heard failure prophesied so oft, been writ\\
\vin So many times among ``The Band''---to wit,\\
The knights who to the Dark Tower's search addressed\\
Their steps---that just to fail as they, seemed best,\\
\vin And all the doubt was now---should I be fit.\\!

8.\\
So, quiet as despair, I turned from him,\\
\vin That hateful cripple, out of his highway\\
\vin Into the path he pointed. All the day\\
Had been a dreary one at best, and dim\\
Was settling to its close, yet shot one grim\\
\vin Red leer to see the plain catch its estray.\\!

9.\\
For mark! no sooner was I fairly found\\
\vin Pledged to the plain, after a pace or two,\\
\vin Than pausing to throw backward a last view\\
To the safe road, 'twas gone! gray plain all round!\\
Nothing but plain to the horizon's bound.\\
\vin I might go on; nought else remained to do.\\!

10.\\
So, on I went. I think I never saw\\
\vin Such starved ignoble nature; nothing throve:\\
\vin For flowers---as well expect a cedar grove!\\
But cockle, spurge, according to their law\\
Might propagate their kind, with none to awe,\\
\vin You'd think: a burr had been a treasure-trove.\\!

11.\\
No! penury, inertness, and grimace,\\
\vin In some strange sort, were the land's portion. ``See\\
\vin Or shut your eyes,'' ---said Nature peevishly---\\
``It nothing skills: I cannot help my case:\\
The Judgment's fire must cure this place,\\
\vin Calcine its clods and set my prisoners free.''\\!

12.\\
If there pushed any ragged thistle-stalk\\
\vin Above its mates, the head was chopped---the bents\\
\vin Were jealous else. What made those holes and rents\\
In the dock's harsh swarth leaves---bruised as to baulk\\
All hope of greenness? 'tis a brute must walk\\
\vin Pashing their life out, with a brute's intents.\\!

13.\\
As for the grass, it grew as scant as hair\\
\vin In leprosy---thin dry blades pricked the mud\\
\vin Which underneath looked kneaded up with blood.\\
One stiff blind horse, his every bone a-stare,\\
Stood stupefied, however he came there---\\
\vin Thrust out past service from the devil's stud!\\!

14.\\
Alive? he might be dead for all I know,\\
\vin With that red gaunt and colloped neck a-strain,\\
\vin And shut eyes underneath the rusty mane.\\
Seldom went such grotesqueness with such woe:\\
I never saw a brute I hated so---\\
\vin He must be wicked to deserve such pain.\\!

15.\\
I shut my eyes and turned them on my heart.\\
\vin As a man calls for wine before he fights,\\
\vin I asked one draught of earlier, happier sights,\\
Ere fitly I could hope to play my part.\\
Think first, fight afterwards---the soldier's art:\\
\vin One taste of the old time sets all to rights!\\!

16.\\
Not it! I fancied Cuthbert's reddening face\\
\vin Beneath its garniture of curly gold,\\
\vin Dear fellow, till I almost felt him fold\\
An arm in mine to fix me to the place,\\
That way he used. Alas! one night's disgrace!\\
\vin Out went my heart's new fire and left it cold.\\!

17.\\
Giles, then, the soul of honour---there he stands\\
\vin Frank as ten years ago when knighted first.\\
\vin What honest men should dare (he said) he durst.\\
Good---but the scene shifts---faugh! what hangman's\\
Pin to his breast a parchment? His own hands\\
\vin Read it. Poor traitor, spit upon and curst!\\!

18.\\
Better this present than a past like that---\\
\vin Back therefore to my darkening path again.\\
\vin No sound, no sight as far as eye could strain.\\
Will the night send a howlet or a bat?\\
I asked: when something on the dismal flat\\
\vin Came to arrest my thoughts and change their train.\\!

19.\\
A sudden little river crossed my path\\
\vin As unexpected as a serpent comes.\\
\vin No sluggish tide congenial to the glooms---\\
This, as it frothed by, might have been a bath\\
For the fiend's glowing hoof---to see the wrath\\
\vin Of its black eddy bespate with flakes and spumes.\\!

20.\\
So, petty yet so spiteful! all along\\
\vin Low scrubby alders kneeled down over it;\\
\vin Drenched willows flung them headlong in a fit\\
Of mute despair, a suicidal throng:\\
The river which had done them all the wrong,\\
\vin Whate'er that was, rolled by, deterred no whit.\\!

21.\\
Which, while I forded,---good saints, how I feared\\
\vin To set my foot upon a dead man's cheek,\\
\vin Each step, or feel the spear I thrust to seek\\
For hollows, tangled in his hair or beard!\\
---It may have been a water-rat I speared,\\
\vin But, ugh! it sounded like a baby's shriek.\\!

22.\\
Glad was I when I reached the other bank.\\
\vin Now for a better country. Vain presage!\\
\vin Who were the strugglers, what war did they wage\\
Whose savage trample thus could pad the dank\\
Soil to a plash? toads in a poisoned tank,\\
\vin Or wild cats in a redhot iron cage---\\!

23.\\
The fight must so have seemed in that fell cirque.\\
\vin What kept them there, with all the plain to choose?\\
\vin No foot-print leading to that horrid mews,\\
None out of it: mad brewage set to work\\
Their brains, no doubt, like galley-slaves the Turk\\
\vin Pits for his pastime, Christians against Jews.\\!

24.\\
And more than that---a furlong on---why, there!\\
\vin What bad use was that engine for, that wheel,\\
\vin Or brake, not wheel---that harrow fit to reel\\
Men's bodies out like silk? with all the air\\
Of Tophet's tool, on earth left unaware,\\
\vin Or brought to sharpen its rusty teeth of steel.\\!

25.\\
Then came a bit of stubbed ground, once a wood,\\
\vin Next a marsh, it would seem, and now mere earth\\
\vin Desperate and done with; (so a fool finds mirth,\\
Makes a thing and then mars it, till his mood\\
Changes and off he goes!) within a rood\\
\vin Bog, clay and rubble, sand and stark black dearth.\\!

26.\\
Now blotches rankling, coloured gay and grim,\\
\vin Now patches where some leanness of the soil's\\
\vin Broke into moss or substances like boils;\\
Then came some palsied oak, a cleft in him\\
Like a distorted mouth that splits its rim\\
\vin Gaping at death, and dies while it recoils.\\!

27.\\
And just as far as ever from the end!\\
\vin Nought in the distance but the evening, nought,\\
\vin To point my footstep further! At the thought,\\
A great black bird, Apollyon's bosom-friend,\\
Sailed past, nor beat his wide wing dragon-penned\\
\vin That brushed my cap---perchance the guide I sought.\\!

28.\\
For, looking up aware I somehow grew,\\
\vin 'Spite of the dusk, the plain had given place\\
\vin All round to mountains---with such name to grace\\
Mere ugly heights and heaps now stol'n in view.\\
How thus they had surprised me,---solve it, you!\\
\vin How to get from them was no plainer case.\\!

29.\\
Yet half I seemed to recognise some trick\\
\vin Of mischief happened to me, God knows when---\\
\vin In a bad dream perhaps. Here ended, then,\\
Progress this way. When, in the very nick\\
Of giving up, one time more, came a click\\
\vin As when a trap shuts---you're inside the den!\\!

30.\\
Burningly it came on me all at once,\\
\vin This was the place! those two hills on the right\\
\vin Crouched like two bulls locked horn in horn in fight---\\
While to the left, a tall scalped mountain . . . Dunce,\\
Fool, to be dozing at the very nonce,\\
\vin After a life spent training for the sight!\\!

31.\\
What in the midst lay but the Tower itself?\\
\vin The round squat turret, blind as the fool's heart,\\
\vin Built of brown stone, without a counterpart\\
In the whole world. The tempest's mocking elf\\
Points to the shipman thus the unseen shelf\\
\vin He strikes on, only when the timbers start.\\!

32.\\
Not see? because of night perhaps?---Why, day\\
\vin Came back again for that! before it left,\\
\vin The dying sunset kindled through a cleft:\\
The hills, like giants at a hunting, lay---\\
Chin upon hand, to see the game at bay,---\\
\vin ``Now stab and end the creature---to the heft!''\\!

33.\\
Not hear? when noise was everywhere? it tolled\\
\vin Increasing like a bell. Names in my ears,\\
\vin Of all the lost adventurers my peers,---\\
How such an one was strong, and such was bold,\\
And such was fortunate, yet each of old\\
\vin Lost, lost! one moment knelled the woe of years.\\!

34.\\
There they stood, ranged along the hill-sides---met\\
\vin To view the last of me, a living frame\\
\vin For one more picture! in a sheet of flame\\
I saw them and I knew them all. And yet\\
Dauntless the slug-horn to my lips I set\\
\vin And blew. ``Childe Roland to the Dark Tower came.''
\end{verse}

\clearpage

% ── Poem 3 ──────────────────────────────────────────────────
\poemtitle{The Tyger}
\poemauthor{William Blake}

\begin{verse}
Tyger Tyger, burning bright,\\
In the forests of the night;\\
What immortal hand or eye,\\
Could frame thy fearful symmetry?\\!

In what distant deeps or skies.\\
Burnt the fire of thine eyes?\\
On what wings dare he aspire?\\
What the hand, dare seize the fire?\\!

And what shoulder, \& what art,\\
Could twist the sinews of thy heart?\\
And when thy heart began to beat.\\
What dread hand? \& what dread feet?\\!

What the hammer? what the chain,\\
In what furnace was thy brain?\\
What the anvil? what dread grasp.\\
Dare its deadly terrors clasp?\\!

When the stars threw down their spears\\
And water'd heaven with their tears:\\
Did he smile his work to see?\\
Did he who made the Lamb make thee?\\!

Tyger Tyger burning bright,\\
In the forests of the night:\\
What immortal hand or eye,\\
Dare frame thy fearful symmetry?
\end{verse}

\clearpage

% ── Poem 4 ──────────────────────────────────────────────────
\poemtitle{``Tomorrow, and tomorrow, and tomorrow''}
\poemauthor{William Shakespeare (from \textit{Macbeth})}

\begin{verse}
Tomorrow, and tomorrow, and tomorrow,\\
Creeps in this petty pace from day to day,\\
To the last syllable of recorded time;\\
And all our yesterdays have lighted fools\\
The way to dusty death. Out, out, brief candle!\\
Life's but a walking shadow, a poor player,\\
That struts and frets his hour upon the stage,\\
And then is heard no more. It is a tale\\
Told by an idiot, full of sound and fury,\\
Signifying nothing.
\end{verse}

\clearpage

% ── Poem 5 ──────────────────────────────────────────────────
\poemtitle{``To be, or not to be, that is the question''}
\poemauthor{William Shakespeare (from \textit{Hamlet})}

\begin{verse}
To be, or not to be, that is the question:\\
Whether 'tis nobler in the mind to suffer\\
The slings and arrows of outrageous fortune,\\
Or to take arms against a sea of troubles\\
And by opposing end them. To die---to sleep,\\
No more; and by a sleep to say we end\\
The heart-ache and the thousand natural shocks\\
That flesh is heir to: 'tis a consummation\\
Devoutly to be wish'd. To die, to sleep;\\
To sleep, perchance to dream---ay, there's the rub:\\
For in that sleep of death what dreams may come,\\
When we have shuffled off this mortal coil,\\
Must give us pause---there's the respect\\
That makes calamity of so long life.\\
For who would bear the whips and scorns of time,\\
Th'oppressor's wrong, the proud man's contumely,\\
The pangs of dispriz'd love, the law's delay,\\
The insolence of office, and the spurns\\
That patient merit of th'unworthy takes,\\
When he himself might his quietus make\\
With a bare bodkin? Who would fardels bear,\\
To grunt and sweat under a weary life,\\
But that the dread of something after death,\\
The undiscovere'd country, from whose bourn\\
No traveller returns, puzzles the will,\\
And makes us rather bear those ills we have\\
Than fly to others that we know not of?\\
Thus conscience doth make cowards of us all,\\
And thus the native hue of resolution\\
Is sicklied o'er with the pale cast of thought,\\
And enterprises of great pith and moment\\
With this regard their currents turn awry\\
And lose the name of action.
\end{verse}

\clearpage

% ── Poem 6 ──────────────────────────────────────────────────
\poemtitle{``Get thee to a nunnery''}
\poemauthor{William Shakespeare (from \textit{Hamlet})}

\begin{verse}
Get thee to a nunnery. Why wouldst thou be\\
a breeder of sinners? I am myself indifferent honest,\\
but yet I could accuse me of such things that it\\
were better my mother had not borne me: I am\\
very proud, revengeful, ambitious, with more offenses\\
at my beck than I have thoughts to put them\\
in, imagination to give them shape, or time to act\\
them in. What should such fellows as I do crawling\\
between earth and heaven? We are arrant knaves\\
all; believe none of us. Go thy ways to a nunnery.\\
Where's your father?
\end{verse}

% ── Poem 7 ──────────────────────────────────────────────────
\poemtitle{She Walks in Beauty}
\poemauthor{Lord Byron (George Gordon)}

\begin{verse}
She walks in beauty, like the night\\
Of cloudless climes and starry skies;\\
And all that's best of dark and bright\\
Meet in her aspect and her eyes;\\
Thus mellowed to that tender light\\
Which heaven to gaudy day denies.\\!

One shade the more, one ray the less,\\
Had half impaired the nameless grace\\
Which waves in every raven tress,\\
Or softly lightens o'er her face;\\
Where thoughts serenely sweet express,\\
How pure, how dear their dwelling-place.\\!

And on that cheek, and o'er that brow,\\
So soft, so calm, yet eloquent,\\
The smiles that win, the tints that glow,\\
But tell of days in goodness spent,\\
A mind at peace with all below,\\
A heart whose love is innocent!
\end{verse}

% ── Poem 8 ──────────────────────────────────────────────────
\poemtitle{Ozymandias}
\poemauthor{Percy Bysshe Shelley}

\begin{verse}
I met a traveller from an antique land\\
Who said: Two vast and trunkless legs of stone\\
Stand in the desert. Near them, on the sand,\\
Half sunk, a shattered visage lies, whose frown,\\
And wrinkled lip, and sneer of cold command,\\
Tell that its sculptor well those passions read\\
Which yet survive, stamped on these lifeless things,\\
The hand that mocked them and the heart that fed:\\
And on the pedestal these words appear:\\
``My name is Ozymandias, king of kings:\\
Look on my works, ye Mighty, and despair!''\\
Nothing beside remains. Round the decay\\
Of that colossal wreck, boundless and bare\\
The lone and level sands stretch far away.
\end{verse}

% ── Poem 9 ──────────────────────────────────────────────────
\poemtitle{``Friends, Romans, countrymen, lend me your ears''}
\poemauthor{William Shakespeare (from \textit{Julius Caesar}, spoken by Marc Antony)}

\begin{verse}
Friends, Romans, countrymen, lend me your ears;\\
I come to bury Caesar, not to praise him.\\
The evil that men do lives after them;\\
The good is oft interred with their bones;\\
So let it be with Caesar. The noble Brutus\\
Hath told you Caesar was ambitious:\\
If it were so, it was a grievous fault,\\
And grievously hath Caesar answer'd it.\\
Here, under leave of Brutus and the rest--\\
For Brutus is an honourable man;\\
So are they all, all honourable men--\\
Come I to speak in Caesar's funeral.\\
He was my friend, faithful and just to me:\\
But Brutus says he was ambitious;\\
And Brutus is an honourable man.\\
He hath brought many captives home to Rome\\
Whose ransoms did the general coffers fill:\\
Did this in Caesar seem ambitious?\\
When that the poor have cried, Caesar hath wept:\\
Ambition should be made of sterner stuff:\\
Yet Brutus says he was ambitious;\\
And Brutus is an honourable man.\\
You all did see that on the Lupercal\\
I thrice presented him a kingly crown,\\
Which he did thrice refuse: was this ambition?\\
Yet Brutus says he was ambitious;\\
And, sure, he is an honourable man.\\
I speak not to disprove what Brutus spoke,\\
But here I am to speak what I do know.\\
You all did love him once, not without cause:\\
What cause withholds you then, to mourn for him?\\
O judgment! thou art fled to brutish beasts,\\
And men have lost their reason. Bear with me;\\
My heart is in the coffin there with Caesar,\\
And I must pause till it come back to me.
\end{verse}

% ── Poem 10 ──────────────────────────────────────────────────
\poemtitle{Lovers and Madmen}
\poemauthor{William Shakespeare (from \textit{A Midsummer Night's Dream}, spoken by Theseus)}

\begin{verse}
More strange than true. I never may believe\\
These antique fables nor these fairy toys.\\
Lovers and madmen have such seething brains,\\
Such shaping fantasies, that apprehend\\
More than cool reason ever comprehends.\\
The lunatic, the lover, and the poet\\
Are of imagination all compact.\\
One sees more devils than vast hell can hold:\\
That is the madman. The lover, all as frantic,\\
Sees Helen's beauty in a brow of Egypt.\\
The poet's eye, in a fine frenzy rolling,\\
Doth glance from heaven to Earth, from Earth to heaven,\\
And as imagination bodies forth\\
The forms of things unknown, the poet's pen\\
Turns them to shapes and gives to airy nothing\\
A local habitation and a name.\\
Such tricks hath strong imagination\\
That, if it would but apprehend some joy,\\
It comprehends some bringer of that joy.\\
Or in the night, imagining some fear,\\
How easy is a bush supposed a bear!
\end{verse}

% ── Poem 11 ──────────────────────────────────────────────────
\poemtitle{``Once more unto the breach, dear friends, once more''}
\poemauthor{William Shakespeare (from \textit{Henry V}, spoken by King Henry)}

\begin{verse}
Once more unto the breach, dear friends, once more;\\
Or close the wall up with our English dead.\\
In peace there's nothing so becomes a man\\
As modest stillness and humility:\\
But when the blast of war blows in our ears,\\
Then imitate the action of the tiger;\\
Stiffen the sinews, summon up the blood,\\
Disguise fair nature with hard-favour'd rage;\\
Then lend the eye a terrible aspect;\\
Let pry through the portage of the head\\
Like the brass cannon; let the brow o'erwhelm it\\
As fearfully as doth a galled rock\\
O'erhang and jutty his confounded base,\\
Swill'd with the wild and wasteful ocean.\\
Now set the teeth and stretch the nostril wide,\\
Hold hard the breath and bend up every spirit\\
To his full height. On, on, you noblest English.\\
Whose blood is fet from fathers of war-proof!\\
Fathers that, like so many Alexanders,\\
Have in these parts from morn till even fought\\
And sheathed their swords for lack of argument:\\
Dishonour not your mothers; now attest\\
That those whom you call'd fathers did beget you.\\
Be copy now to men of grosser blood,\\
And teach them how to war. And you, good yeoman,\\
Whose limbs were made in England, show us here\\
The mettle of your pasture; let us swear\\
That you are worth your breeding; which I doubt not;\\
For there is none of you so mean and base,\\
That hath not noble lustre in your eyes.\\
I see you stand like greyhounds in the slips,\\
Straining upon the start. The game's afoot:\\
Follow your spirit, and upon this charge\\
Cry 'God for Harry, England, and Saint George!'
\end{verse}

\clearpage

% ── Poem 12 ──────────────────────────────────────────────────
\poemtitle{The Lotos-eaters}
\poemauthor{Alfred, Lord Tennyson}

\begin{verse}
"Courage!" he said, and pointed toward the land,\\
"This mounting wave will roll us shoreward soon."\\
In the afternoon they came unto a land\\
In which it seemed always afternoon.\\
All round the coast the languid air did swoon,\\
Breathing like one that hath a weary dream.\\
Full-faced above the valley stood the moon;\\
And like a downward smoke, the slender stream\\
Along the cliff to fall and pause and fall did seem.\\!

A land of streams! some, like a downward smoke,\\
Slow-dropping veils of thinnest lawn, did go;\\
And some thro' wavering lights and shadows broke,\\
Rolling a slumbrous sheet of foam below.\\
They saw the gleaming river seaward flow\\
From the inner land: far off, three mountain-tops,\\
Three silent pinnacles of aged snow,\\
Stood sunset-flush'd: and, dew'd with showery drops,\\
Up-clomb the shadowy pine above the woven copse.\\!

The charmed sunset linger'd low adown\\
In the red West: thro' mountain clefts the dale\\
Was seen far inland, and the yellow down\\
Border'd with palm, and many a winding vale\\
And meadow, set with slender galingale;\\
A land where all things always seem'd the same!\\
And round about the keel with faces pale,\\
Dark faces pale against that rosy flame,\\
The mild-eyed melancholy Lotos-eaters came.\\!

Branches they bore of that enchanted stem,\\
Laden with flower and fruit, whereof they gave\\
To each, but whoso did receive of them,\\
And taste, to him the gushing of the wave\\
Far far away did seem to mourn and rave\\
On alien shores; and if his fellow spake,\\
His voice was thin, as voices from the grave;\\
And deep-asleep he seem'd, yet all awake,\\
And music in his ears his beating heart did make.\\!

They sat them down upon the yellow sand,\\
Between the sun and moon upon the shore;\\
And sweet it was to dream of Fatherland,\\
Of child, and wife, and slave; but evermore\\
Most weary seem'd the sea, weary the oar,\\
Weary the wandering fields of barren foam.\\
Then some one said, "We will return no more";\\
And all at once they sang, "Our island home\\
Is far beyond the wave; we will no longer roam."\\!

\textsc{Choric Song}\\!

I\\
There is sweet music here that softer falls\\
Than petals from blown roses on the grass,\\
Or night-dews on still waters between walls\\
Of shadowy granite, in a gleaming pass;\\
Music that gentlier on the spirit lies,\\
Than tir'd eyelids upon tir'd eyes;\\
Music that brings sweet sleep down from the blissful skies.\\
Here are cool mosses deep,\\
And thro' the moss the ivies creep,\\
And in the stream the long-leaved flowers weep,\\
And from the craggy ledge the poppy hangs in sleep."\\!

II\\
Why are we weigh'd upon with heaviness,\\
And utterly consumed with sharp distress,\\
While all things else have rest from weariness?\\
All things have rest: why should we toil alone,\\
We only toil, who are the first of things,\\
And make perpetual moan,\\
Still from one sorrow to another thrown:\\
Nor ever fold our wings,\\
And cease from wanderings,\\
Nor steep our brows in slumber's holy balm;\\
Nor harken what the inner spirit sings,\\
"There is no joy but calm!"\\
Why should we only toil, the roof and crown of things?\\!

III\\
Lo! in the middle of the wood,\\
The folded leaf is woo'd from out the bud\\
With winds upon the branch, and there\\
Grows green and broad, and takes no care,\\
Sun-steep'd at noon, and in the moon\\
Nightly dew-fed; and turning yellow\\
Falls, and floats adown the air.\\
Lo! sweeten'd with the summer light,\\
The full-juiced apple, waxing over-mellow,\\
Drops in a silent autumn night.\\
All its allotted length of days\\
The flower ripens in its place,\\
Ripens and fades, and falls, and hath no toil,\\
Fast-rooted in the fruitful soil.\\!

IV\\
Hateful is the dark-blue sky,\\
Vaulted o'er the dark-blue sea.\\
Death is the end of life; ah, why\\
Should life all labour be?\\
Let us alone. Time driveth onward fast,\\
And in a little while our lips are dumb.\\
Let us alone. What is it that will last?\\
All things are taken from us, and become\\
Portions and parcels of the dreadful past.\\
Let us alone. What pleasure can we have\\
To war with evil? Is there any peace\\
In ever climbing up the climbing wave?\\
All things have rest, and ripen toward the grave\\
In silence; ripen, fall and cease:\\
Give us long rest or death, dark death, or dreamful ease.\\!

V\\
How sweet it were, hearing the downward stream,\\
With half-shut eyes ever to seem\\
Falling asleep in a half-dream!\\
To dream and dream, like yonder amber light,\\
Which will not leave the myrrh-bush on the height;\\
To hear each other's whisper'd speech;\\
Eating the Lotos day by day,\\
To watch the crisping ripples on the beach,\\
And tender curving lines of creamy spray;\\
To lend our hearts and spirits wholly\\
To the influence of mild-minded melancholy;\\
To muse and brood and live again in memory,\\
With those old faces of our infancy\\
Heap'd over with a mound of grass,\\
Two handfuls of white dust, shut in an urn of brass!\\!

VI\\
Dear is the memory of our wedded lives,\\
And dear the last embraces of our wives\\
And their warm tears: but all hath suffer'd change:\\
For surely now our household hearths are cold,\\
Our sons inherit us: our looks are strange:\\
And we should come like ghosts to trouble joy.\\
Or else the island princes over-bold\\
Have eat our substance, and the minstrel sings\\
Before them of the ten years' war in Troy,\\
And our great deeds, as half-forgotten things.\\
Is there confusion in the little isle?\\
Let what is broken so remain.\\
The Gods are hard to reconcile:\\
'Tis hard to settle order once again.\\
There is confusion worse than death,\\
Trouble on trouble, pain on pain,\\
Long labour unto aged breath,\\
Sore task to hearts worn out by many wars\\
And eyes grown dim with gazing on the pilot-stars.\\!

VII\\
But, propt on beds of amaranth and moly,\\
How sweet (while warm airs lull us, blowing lowly)\\
With half-dropt eyelid still,\\
Beneath a heaven dark and holy,\\
To watch the long bright river drawing slowly\\
His waters from the purple hill---\\
To hear the dewy echoes calling\\
From cave to cave thro' the thick-twined vine---\\
To watch the emerald-colour'd water falling\\
Thro' many a wov'n acanthus-wreath divine!\\
Only to hear and see the far-off sparkling brine,\\
Only to hear were sweet, stretch'd out beneath the pine.\\!

VIII\\
The Lotos blooms below the barren peak:\\
The Lotos blows by every winding creek:\\
All day the wind breathes low with mellower tone:\\
Thro' every hollow cave and alley lone\\
Round and round the spicy downs the yellow Lotos-dust is blown.\\
We have had enough of action, and of motion we,\\
Roll'd to starboard, roll'd to larboard, when the surge was seething free,\\
Where the wallowing monster spouted his foam-fountains in the sea.\\
Let us swear an oath, and keep it with an equal mind,\\
In the hollow Lotos-land to live and lie reclined\\
On the hills like Gods together, careless of mankind.\\
For they lie beside their nectar, and the bolts are hurl'd\\
Far below them in the valleys, and the clouds are lightly curl'd\\
Round their golden houses, girdled with the gleaming world:\\
Where they smile in secret, looking over wasted lands,\\
Blight and famine, plague and earthquake, roaring deeps and fiery sands,\\
Clanging fights, and flaming towns, and sinking ships, and praying hands.\\
But they smile, they find a music centred in a doleful song\\
Steaming up, a lamentation and an ancient tale of wrong,\\
Like a tale of little meaning tho' the words are strong;\\
Chanted from an ill-used race of men that cleave the soil,\\
Sow the seed, and reap the harvest with enduring toil,\\
Storing yearly little dues of wheat, and wine and oil;\\
Till they perish and they suffer---some, 'tis whisper'd---down in hell\\
Suffer endless anguish, others in Elysian valleys dwell,\\
Resting weary limbs at last on beds of asphodel.\\
Surely, surely, slumber is more sweet than toil, the shore\\
Than labour in the deep mid-ocean, wind and wave and oar;\\
O, rest ye, brother mariners, we will not wander more.
\end{verse}

\end{document}
